Lightweight Real-Time Feature Monitoring

### 2020-07-20 paper.tex
Comment out synthetic data experiments
Final paper

\section{Related Work}
\label{sec:RelatedWork}

@@

            \ForEach {$count \in counts$} \Comment{//EMA suppression}
                \State $count \gets count \times 2^{- \frac{1}{n_{1/2}}}$
            \EndFor
        \EndFunction
    \end{algorithmic}
\end{algorithm}
This procedure is illustrated in Figure \ref{fig:EMA-hist-build}.
\begin{figure}[!htb]
    \begin{center}
    %\hspace{6cm}
      \includegraphics[scale=0.6]{figures/ema-hist.png}
      \caption[]{Reference equal-height histogram}
      \label{fig:EMA-hist-build}
    \end{center}
\end{figure}
The EMA-based approximate histogram emulates a sliding window as an exponential tail, seen in Figure \ref{fig:EMA-tail}. This aggregation shows constant memory and time complexity regarding window size, making it ideal for our lightweight real-time system.
\begin{figure}[!htb]
    \begin{center}
    %\hspace{6cm}
      \includegraphics[scale=0.35]{figures/emavswindow.pdf}
      \caption[]{EMA tail versus sliding window weights}
      \label{fig:EMA-tail}
    \end{center}
\end{figure}

@@

\section{Experiments with Synthetic Data}
\label{sec:SyntheticExperiments}
In order to test our proposed system, we generated multiple synthetic datasets before moving on to real data.

@@

\section{Experiments with Real Data}

Due to time constraints, we experimented with data from only one merchant. The merchant's entity as well as feature names will remain anonymous due to confidentiality reasons.

Feedzai provides a fraud detection solution for this merchant and had stored two contiguous in time datasets of credit card transactions that, combined, covered a period of roughly 1 year and 9 months. In this experiment, we used the older one as the reference dataset and the more recent one as the target period. Both datasets were already feature-engineered \cite{Domingos-ML-Feat-Eng} and had a total of 183 numerical plus numerically encoded categorical features (\textit{e.g.}, one-hot encoded or ordinal encoded \cite{categoricalencoding}). Table \ref{tbl:merchant1-datasets-summary} shows summary statistics for both datasets.
\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|c|}
        \hline
        \textbf{Dataset} & \textbf{From}              & \textbf{To}           & \multicolumn{1}{l|}{\textbf{Transactions}} \\ \hline
        Reference        & September 5, 2017 & June 29, 2018 & 1,604,509                                  \\ \hline
        Target           & June 29, 2018      & June 30, 2019 & 4,032,505                                  \\ \hline
        \end{tabular}
        \caption{Summary statistics for reference and target merchant datasets}
        \label{tbl:merchant1-datasets-summary}
    \end{center}
\end{table}

@@

Feature $x_1$ has less alerts in this experiment where we use equal reference and target periods. However, the alarms last for about two months. This is unexpected as we predicted we would see no alarms or few ones with short duration. Likewise, we observe that for feature $x_2$ (Figure \ref{fig:merchant2-x2-correctedpvalues}) and $x_3$ (Figure \ref{fig:merchant2-x3-correctedpvalues}) we still have many alarms and some lasting up to one month.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant2-x1-correctedpvalues.pdf}
      \caption{Merchant feature $x_1$ corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x1-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.25]{figures/merchant2-x2-correctedpvalues.pdf}
      \caption{Merchant feature $x_2$ corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x2-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.25]{figures/merchant2-x3-correctedpvalues.pdf}
      \caption{Merchant feature $x_3$ corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x3-correctedpvalues}
    \end{center}
\end{figure}
We do not have a clear explanation for this and due to time constraints we leave further tests to future work. We suspect the issue may be that there is a difference between the batch and streaming phases. In batch we have subsets of events and compute EMA-like histogram aggregations based on those. In streaming, however, we have a continuous stream of data and never expire the 6\% of data that in batch is not present, because it is a finite window.

@@

\section{Discussion} \label{sec:Discussion}
We began by performing single-feature analysis of synthetic data. The results obtained with this set of experiments were what we expected, in the sense that the artificially introduced anomalies in the datasets were all correctly reported. Still working with synthetic data, we performed our first multi-feature analysis. In this scenario, we applied the Holm-Bonferroni multiple test correction to the p-values of each hypothesis. Once again, each feature was correctly reported as divergent at the timestamps where we changed the underlying generating distribution.

We then moved on to real data. We experimented only with data from one merchant. In our first experiment with real data, we ran our method with what we thought was an ideal set of parameters and produced a lot of alarms. A close analysis of the reference and target time-series of each feature revealed that some alarms indeed made sense, as these datasets contained a lot of drift.

>>>>

%\section{Related Work} \label{sec:RelatedWork}

@@

            \ForEach {$count \in counts$} \Comment{//EMA suppression}
                \State $count \gets count \times 2^{- \frac{1}{n_{1/2}}}$
            \EndFor
        \EndFunction
    \end{algorithmic}
\end{algorithm}
This procedure is illustrated in Figure \ref{fig:EMA-hist-build}.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.6]{figures/ema-hist-paper.png}
      \caption[]{Reference equal-height histogram}
      \label{fig:EMA-hist-build}
    \end{center}
\end{figure}
The EMA-based approximate histogram emulates a sliding window as an exponential tail, seen in Figure \ref{fig:EMA-tail}. This aggregation shows constant memory and time complexity regarding window size, making it ideal for our lightweight real-time system.
\begin{figure}[!htb]
    \begin{center}
    %\hspace{6cm}
      \includegraphics[scale=0.35]{figures/emavswindow.pdf}
      \caption[]{EMA tail versus sliding window weights}
      \label{fig:EMA-tail}
    \end{center}
\end{figure}

@@

\iffalse
\section{Experiments with Synthetic Data}
\label{sec:SyntheticExperiments}
In order to test our proposed system, we generated multiple synthetic datasets before moving on to real data.

@@

\fi

\section{Experimental Results}
We experimented with data from only one merchant and its name as well as feature names will remain anonymous due to confidentiality reasons. Feedzai provides a fraud detection solution for this merchant and had stored two contiguous in time datasets of credit card transactions that, combined, covered a period of roughly 1 year and 9 months. In this experiment, we used the older one as the reference dataset and the more recent one as the target period. Both datasets were already feature-engineered \cite{Domingos-ML-Feat-Eng} and had a total of 183 numerical plus numerically encoded categorical features (\textit{e.g.}, one-hot encoded or ordinal encoded \cite{categoricalencoding}). Table \ref{tbl:merchant1-datasets-summary} shows summary statistics for both datasets.
\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|c|}
        \hline
        \textbf{Dataset} & \textbf{From}              & \textbf{To}           & \multicolumn{1}{l|}{\textbf{Transactions}} \\ \hline
        Reference        & September 5, 2017 & June 29, 2018 & 1,604,509                                  \\ \hline
        Target           & June 29, 2018      & June 30, 2019 & 4,032,505                                  \\ \hline
        \end{tabular}
        \caption{Summary statistics for reference and target merchant datasets}
        \label{tbl:merchant1-datasets-summary}
    \end{center}
\end{table}

@@

Feature $x_1$ has less alerts in this experiment where we use equal reference and target periods. However, the alarms last for about two months. This is unexpected as we predicted we would see no alarms or few ones with short duration. Likewise, we observe that for feature $x_2$ (Figure \ref{fig:merchant2-x2-correctedpvalues}) and $x_3$ (Figure \ref{fig:merchant2-x3-correctedpvalues}) we still have many alarms and some lasting up to one month.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.25]{figures/merchant2-x1-correctedpvalues.pdf}
      \caption{Merchant feature $x_1$ corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x1-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.25]{figures/merchant2-x2-correctedpvalues.pdf}
      \caption{Merchant feature $x_2$ corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x2-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.25]{figures/merchant2-x3-correctedpvalues.pdf}
      \caption{Merchant feature $x_3$ corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x3-correctedpvalues}
    \end{center}
\end{figure}
We do not have a clear explanation for this and due to time constraints we leave further tests to future work. We suspect the issue may be that there is a difference between the batch and streaming phases. In batch we have subsets of events and compute EMA-like histogram aggregations based on those. In streaming, however, we have a continuous stream of data and never expire the 6\% of data that in batch is not present, because it is a finite window.

@@

\section{Discussion} \label{sec:Discussion}
In our first experiment, we ran our method with what we thought was an ideal set of parameters and produced a lot of alarms. A close analysis of the reference and target time-series of each feature revealed that some alarms indeed made sense, as these datasets contained a lot of drift.
